---
title: "Visualise fresh weight"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## https://github.com/patricksnelgar/kiwifruit-graph-theory/blob/master/code/exploration.Rmd

library(readr)
library(dplyr)
library(magrittr)
library(here)
library(lubridate)
library(lattice)
library(lme4)
library(insight)


# relational information for each of the quadrants
quadrant_info <- 
	data.frame(Quadrant = 1:36,
			   OffsetX = rep(c(1000, 500, 0, 0, 500, 1000), each = 6),
			   OffsetY = rep(c(2000, 1000, 0, 0, 1000, 2000), 6),
			   MultiplierX = rep(c(-1,1), each = 18),
			   MultiplierY = rep(rep(c(-1,1), each = 3), 6),
			   QuadrantFromLeader = rep(c(3:1, 1:3), 6), 
			   QuadrantFromTrunk = rep(c(3:1, 1:3), each = 6),
			   NorthSouth = rep(c("S", "N"), each = 18),
			   EastWest = rep(rep(
		   						  c("W", "E"), 
		   						  each = 3
		   						  ),
			   			   	  6)
			   )
# x&y coords for overlaying the quadrant IDs
# only works when vine is plotted with leader running horizontal (swap x and y).
quadrant_labels <- data.frame(x = rep(c(-1250, -750, -250, 250, 750, 1250), each = 6),
							  y = rep(c(2500, 1500, 500, -500, -1500, -2500), 6),
							  label = 1:36)
flowering_dates <- read_csv(here("input/flowering_dates.csv")) %>%
					mutate(FloweringColour = tolower(FloweringColour),
						   FloweringDate = dmy(FloweringDate))

source("../data_import/architecture_data_import_and_merge.R")
source("../data_import/shoot_leaf_area_import_and_plot.R")
source("../data_import/shoot_data_import_and_merge.R")
source("../data_import/fruit_data_import_and_merge.R")

## Add factors
all_fruit_data <- within(all_fruit_data, VineUUID <- factor(VineUUID, levels=unique(VineUUID), labels= paste("Vine", unique(VineUUID))))
```


## Background

Kris Kramer-Walter and Patrick Snelgar have conducted a study of 9 G3 vines from flowering 
fruit harvest, the entire population of fruit were stripped off each vine in 2020, 
fruit were then destructively assessed in the fast lab. Three different canopy
structure treatments (Conventional Spur, Strung) were replicated over three vines.

This data set provides the entire population of fruit weights in 2020, and the underlying
sampling distribution for G3, a control for the Gold programme. From this data set we can
examine the underlying within vine variability of fruit weights;

Tabulating Numbers of fruit per vine assessed by Vine and Treatment;

```{r}
knitr::kable(with(all_fruit_data, table(VineUUID, VineTreatmentNoNumber)))
```

## Exploratory analysis

Visualizing the Fresh weight using histograms, conditioned on each Vine, Treatment, and collapsed over Vines;

```{r}
histogram(~ FreshWeight | VineUUID, data=all_fruit_data, scales=list(alternating=FALSE), between=list(x=0.3, y=0.3),
          breaks=30, layout=c(3,3), panel=function(x, subscripts, groups, ...){
            panel.histogram(x, ...)
            panel.abline(v=90, lty=3, col="red", lwd=1.5)
            thres_frac <- sum(x<=90, na.rm=TRUE) / length(x)
            panel.text(50, 12, paste0(round(100*thres_frac,2), "%"), cex=0.5)
          })

histogram(~ FreshWeight | VineTreatmentNoNumber, data=all_fruit_data,
          aspect=1, main="Collapsing over replicate Vines", 
          scales=list(alternating=FALSE), between=list(x=0.3, y=0.3), breaks=30, layout=c(3,1), panel=function(x, subscripts, groups, ...){
            panel.histogram(x, ...)
            panel.abline(v=90, lty=3, col="red", lwd=1.5)
            thres_frac <- sum(x<=90, na.rm=TRUE) / length(x)
            panel.text(50, 10, paste0(round(100*thres_frac,2), "%"), cex=0.6)
          })

histogram(~ FreshWeight, data=all_fruit_data, scales=list(alternating=FALSE), 
          between=list(x=0.3, y=0.3), breaks=30, main="Collapsing over all Vines",
          panel=function(x, subscripts, groups, ...){
            panel.histogram(x, ...)
            panel.abline(v=90, lty=3, col="red", lwd=1.5)
            thres_frac <- sum(x<=90, na.rm=TRUE) / length(x)
            panel.text(50, 10, paste0(round(100*thres_frac,2), "%"))
          })
```


## Summary statistics

Summary statistics for each vine;

```{r}
sum_vine <- cbind(Vine=levels(all_fruit_data$VineUUID), unname(do.call(rbind.data.frame, with(all_fruit_data, tapply(FreshWeight,  VineUUID, summary)))))
colnames(sum_vine) <- c("Vine", "Min", "1st Qu", "Median","Mean", "3rd Qu", "Max", "NA's")
knitr::kable(sum_vine)
```

Standard deviations within Vines;

```{r}
sd_vine <- with(all_fruit_data, tapply(FreshWeight,  VineUUID, sd, na.rm=TRUE))
knitr::kable(sd_vine,  col.names=c("sd(X)"))
```

The range of standard deviations across the 9 Vines is;

```{r}
range(sd_vine)
```

and the average [RMS](https://en.wikipedia.org/wiki/Root_mean_square) standard deviation is;

```{r}
sqrt(mean(sd_vine^2))
```

These summary statistics appear to be consistent with 
[Kiwifruit fruit size distributions, K. J. McAneney, A. C. Richardson, A. E. Green](https://www.tandfonline.com/doi/pdf/10.1080/01140671.1989.10428047)


The coefficient of variation is a useful metric, a value closer to 0 indicates higher signal to noise ratio.

```{r}
with(all_fruit_data, tapply(FreshWeight,  VineUUID, sd, na.rm=TRUE) / 
       tapply(FreshWeight,  VineUUID, mean, na.rm=TRUE))
```

For different Genotypes within families with smaller fruit sizes, a reasonable assumption to make is that the coefficient of variation would be similar to coefficient of variation values observed above.


## Statistical Power

Using underlying standard deviation information we can estimate statistical power based on a desired difference to detect.

Assuming we have a sample to compare to G3 as a positive control, and the variability is similar, we can estimate power required using a two sided t-test to examine if there is a statistical difference between the samples. Using the empirical median standard deviation which is the standard deviation for at least 50% of the samples;

Sampling n=60;

```{r}
delta <- 1:20
    n <- 60

min_sd <- min(sd_vine)    
rms_sd <- sqrt(mean(sd_vine^2))
max_sd <- max(sd_vine)

delta_min <- lapply(delta, function(x)power.t.test(n=n, sd=min_sd, sig.level=0.05, delta=x, type = "two.sample", alternative="two.sided"))
delta_rms <- lapply(delta, function(x)power.t.test(n=n, sd=rms_sd, sig.level=0.05, delta=x, type = "two.sample", alternative="two.sided"))
delta_max <- lapply(delta, function(x)power.t.test(n=n, sd=max_sd, sig.level=0.05, delta=x, type = "two.sample", alternative="two.sided"))


plot(delta, sapply(delta_rms, function(x)x$power), type="l", col="grey", ylab="Statistical power", main=substitute(paste("n=", n, ", rms ", hat(sigma)==x), list(n=n, x=round(rms_sd, 2))))
lines(delta, sapply(delta_min, function(x)x$power), col="green")
lines(delta, sapply(delta_max, function(x)x$power), col="red")

grid()
legend("topleft",legend = c("max", "rms", "min"), col=c("red","grey", "green"), lty=1, bty="n")
```

Sampling n=40;

```{r}
delta <- 1:20
    n <- 40

min_sd <- min(sd_vine)    
rms_sd <- sqrt(mean(sd_vine^2))
max_sd <- max(sd_vine)

delta_min <- lapply(delta, function(x)power.t.test(n=n, sd=min_sd, sig.level=0.05, delta=x, type = "two.sample", alternative="two.sided"))
delta_rms <- lapply(delta, function(x)power.t.test(n=n, sd=rms_sd, sig.level=0.05, delta=x, type = "two.sample", alternative="two.sided"))
delta_max <- lapply(delta, function(x)power.t.test(n=n, sd=max_sd, sig.level=0.05, delta=x, type = "two.sample", alternative="two.sided"))


plot(delta, sapply(delta_rms, function(x)x$power), type="l", col="grey", ylab="Statistical power", main=substitute(paste("n=", n, ", rms ", hat(sigma)==x), list(n=n, x=round(rms_sd, 2))))
lines(delta, sapply(delta_min, function(x)x$power), col="green")
lines(delta, sapply(delta_max, function(x)x$power), col="red")

grid()
legend("topleft",legend = c("max", "rms", "min"), col=c("red","grey", "green"), lty=1, bty="n")
```

## Estimating variance components

Using lmer to obtain a reml model;

```{r}
m2 <- lmer(FreshWeight ~ VineTreatmentNoNumber + (VineTreatmentNoNumber | VineUUID), data=all_fruit_data)
summary(m2)
```


### Fixed effects

Investigating methods, conventional is slightly better than Spur which is slightly better than Strung.

```{r}
 get_parameters(m2)
```


### Random effects

Variance components between Vines (3 per Treatment level) within Treatments;

```{r}
c(get_variance_intercept(m2), get_variance_slope(m2))
```


Residual standard deviation within Vines is;

```{r}
sqrt(unname(get_variance_residual(m2)))
```


## Sampling Distribution

As we have the entire population of fruit picked and assessed for these 9 Vines,
we have the underlying sampling distribution. This distribution visually appears 
similar to an underlying normal distribution, testing for normality using normal
Q-Q plots conditioned by Vine;

```{r}
qqmath(~ FreshWeight | VineUUID, data=all_fruit_data,
       distribution = function(p) qnorm(p, mean=0, sd=1), main="Normal Q-Q Plot",
       ylab="Sample Quantiles", xlab="Theoretical quantiles",
       scales=list(alternating=FALSE), between=list(x=0.3, y=0.3),
       breaks=30, layout=c(3,3), pch=16, cex=0.5, aspect=1,
      prepanel = prepanel.qqmathline,
       panel = function(x, ...) {
         panel.qqmath(x, ...)
         panel.qqmathline(x, ...)
       })
```

Testing for normality using normal Q-Q plots by Treatment collapsing over replicates;

```{r}
qqmath(~ FreshWeight | VineTreatmentNoNumber, data=all_fruit_data,
       distribution = function(p) qnorm(p, mean=0, sd=1), main="Normal Q-Q Plot",
       ylab="Sample Quantiles", xlab="Theoretical quantiles",
       scales=list(alternating=FALSE), between=list(x=0.3, y=0.3),
       breaks=30, layout=c(3,1), pch=16, cex=0.5, aspect=1,
      prepanel = prepanel.qqmathline,
       panel = function(x, ...) {
         panel.qqmath(x, ...)
         panel.qqmathline(x, ...)
       })
```


Testing for normality using normal Q-Q plots collapsing over all Vines;

```{r}
qqmath(~ FreshWeight, data=all_fruit_data,
       distribution = function(p) qnorm(p, mean=0, sd=1), main="Normal Q-Q Plot",
       ylab="Sample Quantiles", xlab="Theoretical quantiles",
       scales=list(alternating=FALSE), between=list(x=0.3, y=0.3),
       breaks=30, pch=16, cex=0.5, aspect=1,
      prepanel = prepanel.qqmathline,
       panel = function(x, ...) {
         panel.qqmath(x, ...)
         panel.qqmathline(x, ...)
       })
```


## Bootstrapping

```{r}
tmp <- all_fruit_data[all_fruit_data$VineUUID%in%"Vine 2" & !is.na(all_fruit_data$FreshWeight), ]

#sample(tmp$FreshWeight, 40)

hist(replicate(10000,  mean(sample(tmp$FreshWeight, 40))), breaks=30)
hist(replicate(10000,  sd(sample(tmp$FreshWeight, 40))), breaks=30)

qqplot(replicate(10000,  var(sample(tmp$FreshWeight, 40))), rchisq(1000, 40-1))
abline(0,1)


hist(replicate(10000,  mean(sample(tmp$FreshWeight, 60))), breaks=30)
hist(replicate(10000,  sd(sample(tmp$FreshWeight, 60))), breaks=30)

```


```{r scratchpad, echo=FALSE, eval=FALSE}
xyplot( DryMatter ~ FreshWeight | factor(VineUUID), data=all_fruit_data, scales=list(alternating=FALSE), between=list(x=0.3, y=0.3), breaks=30, layout=c(3,3), pch=16, cex=0.3)


summary(lm(Reaction ~ Days, sleepstudy))
summary(lmer(Reaction ~ Days + (Days | Subject), sleepstudy))


xyplot( DryMatter ~ FreshWeightSlice | factor(VineUUID), data=all_fruit_data, scales=list(alternating=FALSE), between=list(x=0.3, y=0.3), breaks=30, layout=c(3,3), pch=16, cex=0.3)


knitr::kable(with(all_fruit_data, tapply(FreshWeight,  VineUUID, length)), col.names=c("N"))


with(all_fruit_data, tapply(FreshWeight,  VineUUID, shapiro.test))

#
## Anova
#
m1 <- aov(FreshWeight ~ VineUUID, data=all_fruit_data)
summary(m1)
sqrt(summary(m1)[[1]][[3]][2])

#
## sleepstudy
#
xyplot(Reaction ~ 1 + Days | Subject , data=sleepstudy, layout=c(9,2),
       panel=function(x,y, subscripts, groups){
         panel.points(x,y)
         panel.lmline(x,y)
       })

#
## Test for normality
#
replicate(10000, shapiro.test(rt(3, 4))$p) %>%  hist
replicate(1000, shapiro.test(sample(all_fruit_data$FreshWeight, 100, replace=TRUE))$p) %>%  hist


#
## Power simulation 
#
## https://github.com/PlantandFoodResearch/RNASeq_community/blob/master/Workshop_March_2018/presentations/day-2/04_session/exercises/simulation.Rmd

n <- 60

## Simulate two sample, one sided
est_power_one <- power.t.test(n=n, sd=20, power=0.8, type="two.sample", alternative="one.sided")


set.seed(42)
pvals <- replicate(100000, {
  X1 <- rnorm(n, 110, 20)
  X2 <- rnorm(n, 110 + est_power_one$delta, 20)

  t.test(X1, X2, alternative="less")$p.value
})

Sens <- sum(pvals <=0.05)/length(pvals)
print(Sens)

## Simulate two sample, two sided
est_power_two <- power.t.test(n=40, sd=20, power=0.8, type="two.sample", alternative="two.sided")

## Simulate H0
set.seed(42)
pvals <- replicate(100000, {
  X1 <- rnorm(40, 110, 20)
  X2 <- rnorm(40, 110 + est_power_two$delta, 20)

  t.test(X1, X2, alternative="two.sided")$p.value
})

Sens <- sum(pvals <=0.05)/length(pvals)
print(Sens)


X <- rnorm(1000, 100, 20)
hist(replicate(10000,  sd(sample(X, 60))), breaks=30)
qqplot(replicate(10000,  var(sample(X, 40))), rchisq(10000, 40-1))
abline(0,1)


ggplot(data.frame(eta=predict(m2,type="link"),pearson=residuals(m2,type="pearson")),
      aes(x=eta,y=pearson)) +
    geom_point() +
    theme_bw()
```