---
title: "Visualise fresh weight"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## https://github.com/patricksnelgar/kiwifruit-graph-theory/blob/master/code/exploration.Rmd

library(readr)
library(dplyr)
library(magrittr)
library(here)
library(lubridate)
library(lattice)
library(lme4)
library(insight)


# relational information for each of the quadrants
quadrant_info <- 
	data.frame(Quadrant = 1:36,
			   OffsetX = rep(c(1000, 500, 0, 0, 500, 1000), each = 6),
			   OffsetY = rep(c(2000, 1000, 0, 0, 1000, 2000), 6),
			   MultiplierX = rep(c(-1,1), each = 18),
			   MultiplierY = rep(rep(c(-1,1), each = 3), 6),
			   QuadrantFromLeader = rep(c(3:1, 1:3), 6), 
			   QuadrantFromTrunk = rep(c(3:1, 1:3), each = 6),
			   NorthSouth = rep(c("S", "N"), each = 18),
			   EastWest = rep(rep(
		   						  c("W", "E"), 
		   						  each = 3
		   						  ),
			   			   	  6)
			   )
# x&y coords for overlaying the quadrant IDs
# only works when vine is plotted with leader running horizontal (swap x and y).
quadrant_labels <- data.frame(x = rep(c(-1250, -750, -250, 250, 750, 1250), each = 6),
							  y = rep(c(2500, 1500, 500, -500, -1500, -2500), 6),
							  label = 1:36)
flowering_dates <- read_csv(here("input/flowering_dates.csv")) %>%
					mutate(FloweringColour = tolower(FloweringColour),
						   FloweringDate = dmy(FloweringDate))

source("../data_import/architecture_data_import_and_merge.R")
source("../data_import/shoot_leaf_area_import_and_plot.R")
source("../data_import/shoot_data_import_and_merge.R")
source("../data_import/fruit_data_import_and_merge.R")

## Add factors
all_fruit_data <- within(all_fruit_data, VineUUID <- factor(VineUUID, levels=unique(VineUUID), labels= paste("Vine", unique(VineUUID))))
```


## Background

Kris Kramer-Walter and Patrick Snelgar have conducted a study of 9 G3 vines from flowering 
fruit harvest, the entire population of fruit were stripped off each vine in 2020, 
fruit were then destructively assessed in the fast lab. Three different canopy
structure treatments (Conventional Spur, Strung) were replicated over three vines.

This dataset provides the entire population of fruit weights in 2020, and the underlying
sampling distribution for G3, a control for the Gold programme. From this dataset we can
examine the underlying within vine variability of fruit weights;

Tabulating Numbers of fruit per vine assessed by Vine and Treatment;

```{r}
knitr::kable(with(all_fruit_data, table(VineUUID, VineTreatmentNoNumber)))
```

## Exploratory analysis

Visualizing the Fresh weight using histograms, conditioned on each Vine, Treatment, and collapsed over Vines;

```{r}
histogram(~ FreshWeight | VineUUID, data=all_fruit_data, scales=list(alternating=FALSE), 
          between=list(x=0.3, y=0.3), breaks=30, layout=c(3,3))

histogram(~ FreshWeight | VineTreatmentNoNumber, data=all_fruit_data,
          aspect=1, main="Collapsing over replicate Vines", 
          scales=list(alternating=FALSE), between=list(x=0.3, y=0.3), breaks=30, layout=c(3,1))

histogram(~ FreshWeight, data=all_fruit_data, scales=list(alternating=FALSE), 
          between=list(x=0.3, y=0.3), breaks=30, main="Collapsing over all Vines")
```


## Summary statistics

Summary statistics for each vine;

```{r}
sum_vine <- cbind(Vine=levels(all_fruit_data$VineUUID), unname(do.call(rbind.data.frame, with(all_fruit_data, tapply(FreshWeight,  VineUUID, summary)))))
colnames(sum_vine) <- c("Vine", "Min", "1st Qu", "Median","Mean", "3rd Qu", "Max", "NA's")
knitr::kable(sum_vine)
```

Standard deviations within Vines;

```{r}
sd_vine <- with(all_fruit_data, tapply(FreshWeight,  VineUUID, sd, na.rm=TRUE))
knitr::kable(sd_vine,  col.names=c("sd(X)"))
```

The range of standard deviations across the 9 Vines is;

```{r}
range(sd_vine)
```

and the average [RMS](https://en.wikipedia.org/wiki/Root_mean_square) standard deviation is;

```{r}
sqrt(mean(sd_vine^2))
```

These summary statistics appear to be consistent with 
[Kiwifruit fruit size distributions, K. J. McAneney, A. C. Richardson, A. E. Green](https://www.tandfonline.com/doi/pdf/10.1080/01140671.1989.10428047)


## Estimating variance components

Using lmer to obtain a reml model;

```{r}
m2 <- lmer(FreshWeight ~ VineTreatmentNoNumber + (VineTreatmentNoNumber | VineUUID), data=all_fruit_data)
summary(m2)
```


### Fixed effects

Investigating methods, conventional is slightly better than Spur which is slightly better than Strung.

```{r}
 get_parameters(m2)
```


### Random effects

Variance components between Vines (3 per Treatment level) within Treatments;

```{r}
c(get_variance_intercept(m2), get_variance_slope(m2))
```


Residual standard deviation within Vines is;

```{r}
sqrt(unname(get_variance_residual(m2)))
```


## Sampling Distribution

As we have the entire population of fruit picked and assessed for these 9 Vines,
we have the underlying sampling distribution. This distribution visually appears 
similar to an underlying normal distrubution, testing for normality using normal
Q-Q plots conditioned by Vine;

```{r}
qqmath(~ FreshWeight | VineUUID, data=all_fruit_data,
       distribution = function(p) qnorm(p, mean=0, sd=1), main="Normal Q-Q Plot",
       ylab="Sample Quantiles", xlab="Theoretical quantiles",
       scales=list(alternating=FALSE), between=list(x=0.3, y=0.3),
       breaks=30, layout=c(3,3), pch=16, cex=0.5, aspect=1,
      prepanel = prepanel.qqmathline,
       panel = function(x, ...) {
         panel.qqmath(x, ...)
         panel.qqmathline(x, ...)
       })
```

Testing for normality using normal Q-Q plots by Treatment collapsing over replicates;

```{r}
qqmath(~ FreshWeight | VineTreatmentNoNumber, data=all_fruit_data,
       distribution = function(p) qnorm(p, mean=0, sd=1), main="Normal Q-Q Plot",
       ylab="Sample Quantiles", xlab="Theoretical quantiles",
       scales=list(alternating=FALSE), between=list(x=0.3, y=0.3),
       breaks=30, layout=c(3,1), pch=16, cex=0.5, aspect=1,
      prepanel = prepanel.qqmathline,
       panel = function(x, ...) {
         panel.qqmath(x, ...)
         panel.qqmathline(x, ...)
       })
```


Testing for normality using normal Q-Q plots collapsing over all Vines;

```{r}
qqmath(~ FreshWeight, data=all_fruit_data,
       distribution = function(p) qnorm(p, mean=0, sd=1), main="Normal Q-Q Plot",
       ylab="Sample Quantiles", xlab="Theoretical quantiles",
       scales=list(alternating=FALSE), between=list(x=0.3, y=0.3),
       breaks=30, pch=16, cex=0.5, aspect=1,
      prepanel = prepanel.qqmathline,
       panel = function(x, ...) {
         panel.qqmath(x, ...)
         panel.qqmathline(x, ...)
       })
```


## Bootstrapping

```{r}
#tmp <- all_fruit_data[all_fruit_data$VineUUID%in%"Vine 1" & !is.na(all_fruit_data$FreshWeight), ]

#sample(tmp$FreshWeight, 40)

hist(replicate(10000,  var(sample(tmp$FreshWeight, 40))), breaks=30)

qqplot(replicate(10000,  var(sample(tmp$FreshWeight, 40))), rchisq(1000, 40-1))
```


```{r scratchpad, echo=FALSE, eval=FALSE}
xyplot( DryMatter ~ FreshWeight | factor(VineUUID), data=all_fruit_data, scales=list(alternating=FALSE), between=list(x=0.3, y=0.3), breaks=30, layout=c(3,3), pch=16, cex=0.3)


summary(lm(Reaction ~ Days, sleepstudy))
summary(lmer(Reaction ~ Days + (Days | Subject), sleepstudy))


xyplot( DryMatter ~ FreshWeightSlice | factor(VineUUID), data=all_fruit_data, scales=list(alternating=FALSE), between=list(x=0.3, y=0.3), breaks=30, layout=c(3,3), pch=16, cex=0.3)


knitr::kable(with(all_fruit_data, tapply(FreshWeight,  VineUUID, length)), col.names=c("N"))


with(all_fruit_data, tapply(FreshWeight,  VineUUID, shapiro.test))

## Anova
m1 <- aov(FreshWeight ~ VineUUID, data=all_fruit_data)
summary(m1)
sqrt(summary(m1)[[1]][[3]][2])


## sleepstudy
xyplot(Reaction ~ 1 + Days | Subject , data=sleepstudy, layout=c(9,2),
       panel=function(x,y, subscripts, groups){
         panel.points(x,y)
         panel.lmline(x,y)
       })

## Test for normality
replicate(10000, shapiro.test(rt(3, 4))$p) %>%  hist
replicate(1000, shapiro.test(sample(all_fruit_data$FreshWeight, 100, replace=TRUE))$p) %>%  hist
```